Hydra battles in Coq

### 2020-10-23 doc/hydras.tex
Writing a chapter on Gamma0

\begin{remark}
The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
 \end{remark}

@@

\begin{Coqsrc}
Definition epsilon0  := [one,zero].

@@

@@

| (* 4 *)
 lt_4 : forall alpha1 alpha2 beta1 beta2 n1 n2 gamma1 gamma2, 
               alpha2 t2< alpha1 ->
               gcons alpha1 beta1 0 zero t2< beta2 ->
               gcons alpha1 beta1 n1 gamma1 t2<
               gcons alpha2 beta2 n2 gamma2

| (* 5 *)
lt_5 : forall alpha1 alpha2 beta1 n1 n2 gamma1 gamma2, 
               alpha2 t2< alpha1 ->
               gcons alpha1 beta1 n1 gamma1 t2<
               gcons alpha2  (gcons alpha1 beta1 0 zero) n2 gamma2

@@

| (* 7 *)
  lt_7 : forall alpha1 beta1 n1   gamma1 gamma2,  gamma1 t2< gamma2 ->
                                      gcons alpha1 beta1 n1 gamma1 t2<
                                      gcons alpha1 beta1 n1 gamma2
where  "o1 t2< o2" := (lt o1 o2): T2_scope.
\end{Coqsrc}

Here are two simple verifications: $\omega$ is less than $\epsilon_0$, whic is less than $\psi(2,1)$.

\begin{Coqsrc}
Goal omega t2< epsilon0.
Proof.
  unfold epsilon0; auto with T2.
Qed.

Goal epsilon0 t2< [2,1].
Proof.
  unfold epsilon0;  apply lt_2.
  - apply lt_6; auto with arith.
  - auto with T2.
Qed.
\end{Coqsrc}

@@

\chapter{Appendices}

>>>>

\begin{remark}
The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
 \end{remark}

@@

\begin{Coqsrc}
Notation "'epsilon0'"  := ([one,zero]) : T2_scope.

@@

\label{sect:t2-lt-def}

@@

| (* 4 *)
 lt_4 : forall alpha1 alpha2 beta1 beta2 n1 n2 gamma1 gamma2, 
               alpha2 t2< alpha1 ->
               [alpha1, beta1] t2< beta2 ->
               gcons alpha1 beta1 n1 gamma1 t2<
               gcons alpha2 beta2 n2 gamma2

| (* 5 *)
lt_5 : forall alpha1 alpha2 beta1 n1 n2 gamma1 gamma2, 
               alpha2 t2< alpha1 ->
               gcons alpha1 beta1 n1 gamma1 t2<
               gcons alpha2  [alpha1, beta1] n2 gamma2

@@

| (* 7 *)
  lt_7 : forall alpha1 beta1 n1   gamma1 gamma2,  gamma1 t2< gamma2 ->
                                      gcons alpha1 beta1 n1 gamma1 t2<
                                      gcons alpha1 beta1 n1 gamma2
where  "o1 t2< o2" := (lt o1 o2): T2_scope.

Hint Constructors lt : T2.
\end{Coqsrc}

\subsection{Examples}
Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.

\subsubsection*{Proof of $0<\epsilon_0$}

\begin{Coqsrc}
Example Ex1: 0 t2< epsilon0.
Proof.  constructor 1. Qed.
\end{Coqsrc}

\subsubsection*{Proof of $\omega<\epsilon_0$}

\begin{Coqsrc}
Example Ex2: omega t2< epsilon0.
Proof. info_auto with T2. (* uses lt_1 and lt_2 *) Qed.
\end{Coqsrc}

\subsubsection*{Proof of $\psi(\omega,8]\times 13+56 < \psi(\omega,8]\times 13+57 $}

\begin{Coqsrc}
Example Ex3: gcons omega 8 12 56 t2<  gcons omega 8 12 57.
Proof.
  constructor 7; constructor 6; auto with arith.
Qed.
\end{Coqsrc}

\subsubsection*{Proof of $\epsilon_0<\psi(2,1)$}

\begin{Coqsrc}
Example Ex4: epsilon0 t2< [2,1].
Proof.
   constructor 2; auto with T2.
   - constructor 6; auto with arith.
Qed.
\end{Coqsrc}

\subsubsection*{Proof of $\psi(2,1)<\psi(2,3)$}

\begin{Coqsrc}
Example Ex5 : [2,1] t2< [2,3].
Proof.
  constructor 3; auto with T2.
  - constructor 6; auto with arith.
Qed.
\end{Coqsrc}

\subsubsection*{Proof of $\psi(1,0)\times 13+ \omega < \psi(0,\psi(2,1))$}

\begin{Coqsrc}
Example Ex6 : gcons 1 0 12 omega t2< [0,[2,1]].
Proof.
  constructor 4.
  - constructor 1.
  - constructor 2.
    + constructor 6; auto with arith.
    + constructor 1.
Qed.
\end{Coqsrc}

\subsubsection*{Proof of $\psi(2,1)\times 43 + \epsilon_0 < \psi(1,\psi(2,1))$}

\begin{Coqsrc}
Example Ex7 : gcons 2 1 42 epsilon0 t2< [1, [2,1]].
Proof.
 constructor 5.
 constructor 6; auto with arith.
Qed.
\end{Coqsrc}

@@

\section{Veblen normal forms}
\begin{definition}
  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
 \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
\end{definition}

\begin{Coqsrc}
Inductive nf : T2 -> Prop :=
| zero_nf : nf zero
| single_nf : forall a b n, nf a ->  nf b -> nf (gcons a b n zero)
| gcons_nf : forall a b n a' b' n' c', 
                      [a', b'] t2< [a, b]  -> 
                      nf a -> nf b -> 
                      nf(gcons a' b' n' c')-> 
                      nf(gcons a b n (gcons a' b' n' c')).
\end{Coqsrc}

Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).

\begin{Coqsrc}
Lemma  nf_finite i : nf (fin (S i)).
Proof.
   constructor 2; auto with T2.
Qed.

Lemma nf_epsilon0 : nf epsilon0.
Proof. constructor 2; auto with T2. Qed.

Example Ex8: nf (gcons 2 1 42 epsilon0).
Proof.
  constructor 3; auto with T2.
  - apply Ex4.
  - apply nf_finite.
  - apply nf_finite.
Qed.

\end{Coqsrc}
\chapter{Appendices}
